Are we learning to build better models at the cost of better decisions?

### 2021-03-11 manuscript/manuscript.Rmd
:pen:

## Forecasts and Proper Scores

<!-- I think this subsection should be renamed. it makes it a bit hard to follow the structure of the larger section. Seems like something like: "Statistical approaches: forecasts and proper scores" idk, but something that more closely aligns with the above paragraph and the second subsection -->

Like many decision problems, the task of setting a sustainable harvest quota appears to hinge on having an accurate forecast: if we can predict to what size the fish stock will increase next year, $X_t+1$, and we know the current stock, $X_t$, then we can safely harvest $X_{t+1} - X_t$.

<!-- it is unclear what you mean by safely harvest?  as in the stock wont crash? or it won't be worse than it is this year? You could imaging that we could safetly harvest more (or only less) than that.-->

Overestimating or underestimating such recruitment will result in over-harvesting or under-harvesting, respectively.

Selecting a model based on forecast skill is also justifiable on theoretical grounds, since it reduces the risk of over-fitting by comparing model predictions to later observations that were not used to estimate the model.

@Gneiting2007 provides a rigorous proof for "proper" scoring criteria, which have the desirable property which no model predicting the distribution of future outcomes, $Q(x)$ can achieve a better average score than the true model $P(x)$. That is, unlike likelihood or other goodness-of-fit scores, it is impossible to overfit when conditioning on a strictly proper score -- since no model model can out-perform the true model. Not that strictly proper scoring rules score _probabilistic forecasts_ and not just point predictions, favoring models which accurately reflect the uncertainty over those which under-estimate it.  These features have made proper scoring rules for probabilistic forecasts a successful and popular approach for addressing model uncertainty in many other areas [@Gneiting2014; @Raftery2016] and a promising tool for evaluation of ecological forecasts [@Dietze2018].

@@

<!-- Inttuition for it's not just this example -->
This phenomenon is not unique to sequential decision problems or optimal control solutions.  For example, Management Strategy Evaluation [@Punt2016] in fisheries seeks to evaluate pre-specified strategies rather solve for the best possible strategy through optimal control.  This approach is well justified -- optimal control techniques such as stochastic dynamic programming illustrated here quickly become intractable for more complex models [@Marescot2013] typically used in fisheries stock assessments.  Constraining the search to only strategies that impose a constant mortality (defined as harvest per unit biomass, $F = H/B$) means we only have to evaluate those $N$ strategies each iteration, not solve a sequential decision problem. Doing so, we would find the best constant-mortality solution under Model 1 performs much worse than the best constant-mortality solution under Model 2. Does this mean MSE is not susceptible to this issue?  No, it does not.  It is just as easy to construct an alternative Model 1 with the same properties of leading to nearly optimal decisions while being rejected by any method of iterative learning, such as forecasting or adaptive management.  Just as the optimal control policy for a Gordon-Schaefer (logistic) model depends only on parameter $K$ [@Reed1979], the optimum constant escapement policy depends only on parameter $r$ [@Schaefer1954]. If Model 1 has a value of $r$ such that it happens to match the best possible constant-mortality solution for the (unobserved) true model, then $K$ can be set arbitrarily high, ensuring the any model selection, forecasting, or adaptive management approach would lead away from Model 1 and towards worse-performing models once again.

@@

<!-- So what? -->
If these unicorns do exist in real world management, then what do we do about them? I believe this is an open question, but that our first step must be to recognize it as such.  We have seen that the problem cannot be resolved by more data, and is not the result of overfitting. Nor is "creating more models" the answer: when we have a model that is good enough to get optimal results, we cannot always insist on more models. We have also seen how not updating our uncertainty estimates, or doing so less frequently, can reveal a unicorn model before it is discounted by further iterations. Because value of information can be negative, we must revise model uncertainty only cautiously.  More importantly, we have seen that decision theory, such as the intuition offered by Fig 4, can help us better understand what features of a model are essential to decision outcomes and what are not.  Building on such understanding, it may possible to identify strategies for learning that are more agnostic to the details of the model that are not important.

<!-- Model free management -->

>>>>

## Statistical approaches: Forecasting under "Proper" Scoring Rules

Like many decision problems, the task of setting a sustainable harvest quota appears to hinge on having an accurate forecast: if we can predict to what size the fish stock will increase next year, $X_t+1$, and we know the current stock, $X_t$, then we can sustainably harvest $X_{t+1} - X_t$ without decreasing the biomass over the long term.  Selecting a model based on forecast skill is also justifiable on theoretical grounds, since it reduces the risk of over-fitting by comparing model predictions to later observations that were not used to estimate the model [@Gneiting2014].

@Gneiting2007 provides a rigorous proof for such "proper" scoring criteria, which have the desirable property which no model predicting the distribution of future outcomes, $Q(x)$ can achieve a better average score than the true model $P(x)$. That is, unlike likelihood or other goodness-of-fit scores, it is impossible to overfit when conditioning on a strictly proper score -- no model model can out-perform the true model. Not that strictly proper scoring rules score _probabilistic forecasts_ and not just point predictions, favoring models which accurately reflect the uncertainty over those which under-estimate it.  These features have made proper scoring rules for probabilistic forecasts a successful and popular approach for addressing model uncertainty in many other areas [@Gneiting2014; @Raftery2016] and a promising tool for evaluation of ecological forecasts [@Dietze2018].

@@

<!-- Inttuition for it's not just this example -->
This phenomenon is not unique to sequential decision problems or optimal control solutions.  For example, Management Strategy Evaluation [@Punt2016] in fisheries seeks to evaluate pre-specified strategies rather solve for the best possible strategy through optimal control.  This approach is well justified -- optimal control techniques such as stochastic dynamic programming illustrated here quickly become intractable for more complex models [@Marescot2013] typically used in fisheries stock assessments.  Constraining the search to only strategies that impose a constant mortality (defined as harvest per unit biomass, $F = H/B$) means we only have to evaluate those $N$ strategies each iteration, not solve a sequential decision problem. Doing so, we would find the best constant-mortality solution under Model 1 performs much worse than the best constant-mortality solution under Model 2. Does this mean MSE is not susceptible to this issue?  No, it does not.  It is just as easy to construct an alternative Model 1 with the same properties of leading to nearly optimal decisions while being rejected by any method of iterative learning, such as forecasting or adaptive management.  Just as the optimal control policy for a Gordon-Schaefer (logistic) model depends only on parameter $K$ [@Reed1979], the optimum constant escapement policy depends only on parameter $r$ [@Schaefer1954]. If Model 1 has a value of $r$ such that it happens to match the best possible constant-mortality solution for the (unobserved) true model, then $K$ can be set arbitrarily high, ensuring the any model selection, forecasting, or adaptive management approach would lead away from Model 1 and towards worse-performing models once again. Many optimization problems share this feature in which the optimal policy depends only on a subset or ratio of model parameters, such that it is usually easy to find a similar Model 1. This example also illustrates the importance of decision constraints -- a model that gives very good outcomes when we are free to vary the havest quota each year (the optimal control problem here) may give very poor results under the constraint of constant mortality, and vice versa.

@@

<!-- So what? -->
If these unicorns do exist in real world management, then what do we do about them? I believe this is an open question, but that our first step must be to recognize it as such.  We have seen that the problem cannot be resolved by more data, and is not the result of overfitting. Nor is "creating more models" the answer: when we have a model that is good enough to get optimal results, we cannot always insist on more models. We have also seen how not updating our uncertainty estimates, or doing so less frequently, can reveal a unicorn model before it is discounted by further iterations. Approaches such as iterative forecasting or adaptive management that can reduce model uncertainty over time remain promising and important techniques, but because value of information can be negative, we must revise model uncertainty only cautiously.  More importantly, we have seen that decision theory, such as the intuition offered by Fig 4, can help us better understand what features of a model are essential to decision outcomes and what are not.  Building on such understanding, it may possible to identify strategies for learning that are more agnostic to the details of the model that are not important, or perhaps not reliant on model-based predictions at all. How that is done is an open question.

<!-- Model free management -->
<!-- For example, a simple approach could gradually increase harvest from 0 as long as the observed recruitment increases relative to previous values.  This corresponds to a local search for the peak growth-rate from above. This strategy is optimal for the model considered here, at least under small noise, but is not robust to growth rates with local optima, nor is it optimal during a possibly-very-long transient period of the search.   -->
